---
title: "ProFuse AGN PSF"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ProFuse AGN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# ProFound + ProFit + ProSpect = ProFuse

Here we show the minimal working example of full multi-band simultaneous **ProFit** and **ProSpect** fitting, this has been nick-named **ProFuse**, but this is not really the official name (yet...).  The idea is that rather than fitting **ProFit** for each image individually, or even with priors informed by earlier fits and/or a forced smooth relationship of the **ProFit** parameters as a function of wavelength (which might look nice, but could well be unphysical) we use **ProSpect** to fit a true star formation and dust attenuation law etc along with certain **ProFit** properties. The end result is that the likelihood is computed at the per band-image level, and the outcome is we compute a separate star formation / metallicity history etc for the bulge and disk components.

Doing all of this is fairly expensive (example below takes about an hour), but it is actually substantially cheaper when using e.g. 9-band photometry (example below) than running normal **ProFit** 9 times and then **ProSpect** twice (on the bulge and disk outputs), and in any case this certainly would not let guarantee the output is physically meaningful.

To keep things feasible (both in terms of coding and fitting) only a restrictive subset of **ProFit** component combinations are possible- basically the most popular and useful subset of bulge and disk fitting (with flexibility of free Sersic bulge and/or disk etc). My experience has been nobody uses the more complicated and flexible profiling options (sad-face) so until these are under serious demand this will probably be the limit of profiling options.

## Minimal Bulge-Disk Example (with fixed de-V bulge and Exp disk)

Set global evaluate:

```{r}
evalglobal = FALSE
```

Load the required packages (**ProTools** basically load all of my packages in one go):

```{r}
library(Rfits)
library(ProFuse)
library(ProTools)
library(ParmOff)
```

Minimal setup information. note we need to know the true redshift of our source (so this will change with every galaxy) and the cutout location.

```{r}
redshift = 0.0447
data('BC03lr')
data('Dale_NormTot')
data('AGN_UnOb_Sparse')
data('Dale_M2L_func')
data('SKIRTOR')
filters=c('u_VST', 'g_VST', 'r_VST', 'i_VST', 'Z_VISTA',
          'Y_VISTA', 'J_VISTA', 'H_VISTA', 'Ks_VISTA')
filtout={}
for(i in filters){filtout=c(filtout,list(approxfun(getfilt(i))))}

loc = c(1200,480)
cut = -299:300

cenwaves = cenwave[match(filters, cenwave$filter),'cenwave']
  
agemax = 13.3e9 - cosdistTravelTime(z=redshift, H0 = 67.8, OmegaM = 0.308)*1e9
```

Cut-out all of our images: 

```{r, eval=evalglobal}
image_list = list(
  u = Rfits_read_image(system.file("extdata", 'MultiBand/u.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  g = Rfits_read_image(system.file("extdata", 'MultiBand/g.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  r = Rfits_read_image(system.file("extdata", 'MultiBand/r.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  i = Rfits_read_image(system.file("extdata", 'MultiBand/i.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  Z = Rfits_read_image(system.file("extdata", 'MultiBand/Z.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  Y = Rfits_read_image(system.file("extdata", 'MultiBand/Y.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  J = Rfits_read_image(system.file("extdata", 'MultiBand/J.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  H = Rfits_read_image(system.file("extdata", 'MultiBand/H.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut],
  Ks = Rfits_read_image(system.file("extdata", 'MultiBand/Ks.fits', package="ProFound"),ext=2)$imDat[loc[1] + cut, loc[2] + cut]
)
```

The reasonably minimal run of MultiBandFound2Fit, which sets everything up for later fitting. Here we are mostly setting the **ProSpect** related inputs (most of the **ProFit** ones are either reasonable to assume, or can be estimated from the **ProFound** outputs). Note this takes a few minutes to run, because we are also fitting the PSF per band by automatically extracting stars from the image and fitting them. If you already have PSFs for some bands you can provide them in *psf_list*, leaving unknown PSF entries as NULL.

```{r eval=evalglobal, message=FALSE, warning=FALSE, results=FALSE}
MF2F = profuseMultiBandFound2Fit(image_list=image_list,
                                    magzero=c(0,0,0,0,30,30,30,30,30),
                                    tolerance=20,
                                    Ncomp = 2.5,
                                    parm_global = c("sersic.xcen1", "sersic.ycen1",
                                                    "sersic.re1", "sersic.re2",
                                                    "sersic.ang2", "sersic.axrat2"),
                                    parm_ProSpect = list(
                                      AGNlum_1 = 42,
                                      mSFR_2 = 0,
                                      mpeak_2 = 10,
                                      mperiod_2 = 0.3,
                                      mskew_2 = 0,
                                      Zfinal_2 = -2,
                                      mSFR_3 = 0,
                                      mpeak_3 = 1,
                                      mperiod_3 = 0.3,
                                      mskew_3 = 0,
                                      Zfinal_3 = -2
                                    ),
                                    logged_ProSpect = c(
                                      TRUE,                              #         AGNlum = 42,
                                      TRUE,                              #         mSFR_2 = 0,
                                      FALSE,                             #         mpeak_2 = 5,
                                      TRUE,                              #        mperiod_2 = 0.3,
                                      FALSE,                             #         mskew_2 = 0,
                                      TRUE,                              #        Zfinal_2 = -2,
                                      TRUE,                              #        mSFR_3 = 0,
                                      FALSE,                             #         mpeak_3 = 5,
                                      TRUE,                              #        mperiod_3 = 0.3,
                                      FALSE,                             #         mskew_3 = 0,
                                      TRUE                               #        Zfinal_3 = -2,
                                    ),
                                    intervals_ProSpect = list(
                                      lo=c(
                                        30,                              #         AGNlum = 42,
                                        -4,                              #       mSFR_2 = 0,
                                        0,                               #       mpeak_2 = 5,
                                        -0.5,                            #          mperiod_2 = 0.3,
                                        -1,                              #        mskew_2 = 0,
                                        -4,                              #        Zfinal_2 = -2,
                                        -4,                              #        mSFR_3 = 0,
                                        0,                               #       mpeak_3 = 5,
                                        -0.5,                            #          mperiod_3 = 0.3,
                                        -1,                              #        mskew_3 = 0,
                                        -4                               #        Zfinal_3 = -2,
                                      ),
                                      hi=c(
                                        45,                              #         AGNlum = 42,
                                        3,                               #       mSFR_2 = 0,
                                        10,                              #        mpeak_2 = 5,
                                        1,                               #       mperiod_2 = 0.3,
                                        1,                               #       mskew_2 = 0,
                                        -1.3,                              #        Zfinal_2 = -2,
                                        3,                               #       mSFR_3 = 0,
                                        10,                              #        mpeak_3 = 5,
                                        1,                               #       mperiod_3 = 0.3,
                                        1,                               #       mskew_3 = 0,
                                        -1.3                              #        Zfinal_3 = -2,
                                      )
                                    ),
                                    data_ProSpect = list(massfunc = massfunc_snorm_trunc,
                                                         speclib = BC03lr,
                                                         Dale = Dale_NormTot,
                                                         AGN_1 = SKIRTOR,
                                                         filtout = filtout,
                                                         z = redshift,
                                                         Z = Zfunc_massmap_lin,
                                                         agemax = agemax,
                                                         #Set no SFR in PSF
                                                         mSFR_1 = 0,
                                                         #Set no ISM dust in bulge
                                                         tau_screen_2 = 0,
                                                         tau_birth_2 = 0.63,
                                                         #Set to Thorne 2020 medians for disk
                                                         tau_screen_3 = 0.16,
                                                         tau_birth_3 = 0.63
                                                         )
                                    )
```

For convenience we can also just pass our *MF2F* object into the *profuseMultiBandDoFit* function. This will still run Highlander (with slightly restricted options available), but it has the advantage that it will properly pass all of the fitting limits through (a non-trivial task with many image and complex combinations of logged parameters).

Check it:

```{r}
profitLikeModel(MF2F$init, MF2F, makeplots=T)
```

And now fit it (this took about 4 hours):

```{r, eval=evalglobal}
highfit = profuseMultiBandDoFit(MF2F=MF2F)
```


And check the final fit:

```{r}
profitLikeModel(highfit$parm, MF2F, makeplots=T)
```

```{r fig.width=5, fig.height=5, dpi=100, eval=evalglobal}
magcurve(massfunc_snorm_trunc(age=x,mSFR=10^highfit$parm["mSFR_2"],mpeak=highfit$parm["mpeak_2"],mperiod=10^highfit$parm["mperiod_2"],
  mskew=highfit$parm["mskew_2"], magemax=agemax/1e9),0,13e9,add=FALSE,col='red', ylim=c(0,4.5),xlab='Age (Yr)', ylab='SFR (Msol / Yr)')
magcurve(massfunc_snorm_trunc(age=x,mSFR=10^highfit$parm["mSFR_3"],mpeak=highfit$parm["mpeak_3"],mperiod=10^highfit$parm["mperiod_3"],
  mskew=highfit$parm["mskew_3"], magemax=agemax/1e9),0,13e9,add=TRUE,col='blue')
legend('topright', legend=c('Bulge', 'Disk'), col=c('red', 'blue'), lty=1)
```
